The Nature and Orbit of the Ophiuchus Stream

### 2015-01-03 tex/ms.tex
Fixed Heather Flewelling's name. Ran the text through a spellchecker and fixed a few typos. Described the galactocentric cartesian system used in Figure 11.

\author{Branimir Sesar\altaffilmark{1,2}}
\author{Jo Bovy\altaffilmark{3,4}}
\author{Edouard J.~Bernard\altaffilmark{5}}
\author{Nelson Caldwell\altaffilmark{6}}
\author{Judith G.~Cohen\altaffilmark{7}}
\author{Morgan Fouesneau\altaffilmark{1}}
\author{Christian I.~Johnson\altaffilmark{6}}
\author{Melissa Ness\altaffilmark{1}}
%\author{Maria Bergemann\altaffilmark{1}}
\author{Annette M.~N.~Ferguson\altaffilmark{5}}
\author{Nicolas F.~Martin\altaffilmark{8,1}}
\author{Hans-Walter Rix\altaffilmark{1}}
\author{Edward F.~Schlafly\altaffilmark{1}}
%\author{builders}
\author{William S.~Burgett\altaffilmark{15}}
\author{Kenneth C.~Chambers\altaffilmark{9}} 
%\author{L.~Denneau\altaffilmark{9}} 
%\author{P.~Draper\altaffilmark{10}} 
\author{H.~Flewelling\altaffilmark{9}} 
%\author{T.~Grav\altaffilmark{11}} 
%\author{J.~N.~Heasley\altaffilmark{9}} 
\author{Klaus W.~Hodapp\altaffilmark{9}} 
%\author{M.~E.~Huber\altaffilmark{9}} 
%\author{R.~Jedicke\altaffilmark{9}} 
\author{Nick Kaiser\altaffilmark{9}} 
%\author{R.-P.~Kudritzki\altaffilmark{9}}
%\author{G.~A.~Luppino\altaffilmark{9}}
%\author{R.~H.~Lupton\altaffilmark{12}}
\author{Eugene A.~Magnier\altaffilmark{9}}
\author{Imants Platais\altaffilmark{11}}
%\author{N.~Metcalfe\altaffilmark{10}}
%\author{D.~G.~Monet\altaffilmark{13}}
%\author{J.~S.~Morgan\altaffilmark{9}}
%\author{P.~M.~Onaka\altaffilmark{9}}
%\author{P.~A.~Price\altaffilmark{12}}
%\author{C.~W.~Stubbs\altaffilmark{14}}
%\author{W.~Sweeney\altaffilmark{9}}
\author{John L.~Tonry\altaffilmark{9}}
%\author{R.~J.~Wainscoat\altaffilmark{9}}
\author{Christopher Waters\altaffilmark{9}}
\author{Rosemary~F.~G.~Wyse\altaffilmark{11}}

@@

We determined stellar parameters from the continuum-normalised, radial
velocity-corrected spectra using the SMH code of \citet{cas14}, which is built
on the MOOG code of \citet{sne73}. Kurucz $\alpha-$enhanced (0.4 dex) model
atmospheres \citep{ck04} and a line list compiled from \citet{fre10} by
\citet{cas14} were used (see Table~5 in the electronic version of the Journal).
First, effective temperatures were calculated from 2MASS photometry and
color-temperature calibrations of \citet{ghb09} and spectroscopic temperatures
were optimized around this value using the SMH code, by removing abundance
trends with line excitation potential. Parameters of $\log g$ and ${\rm [Fe/H]}$
were determined using 14 Fe I and 2-3 Fe II lines to achieve ionization balance,
and microturbulence was calculated by removing abundance trends with reduced
equivalent width of the lines. Estimates of the $\alpha-$enhancement were
obtained using only the few available clean Mg, Ca and Ti (I and II) lines,
which comprised a total of 6-8 lines per star. The abundances were measured from
equivalent widths and the lines we used are listed in Table~5. Because Mg lines
are strong and may be saturated, the values of ${\rm [Mg/Fe]}$ are significantly
different than values of ${\rm [Ca/Fe]}$ and ${\rm [Ti/Fe]}$.

@@

Having tied the positions for each object at one 2MASS and five USNO-B epochs to
the PS1 astrometric reference frame, we can now check for any additional
systematic uncertainties in the calibrated astrometry. We do so using the
leave-one-out cross-validation. One of the six calibrated positions is witheld, 
and a straight line is fitted to the remaining five positions and the PS1
position. The straight line fit (i.e., essentially a proper motion fit,
neglecting the parallax) is then used to predict the position of an object at
the witheld epoch. The difference between the witheld position and the predicted
position is labeled as $\Delta RA$ or $\Delta Decl$.

@@

To efficiently explore the parameter space, we use the \citet{gw10} Affine
Invariant Markov chain Monte Carlo (MCMC) Ensemble sampler as implemented in the
{\em emcee} package\footnote{\url{http://dan.iel.fm/emcee/current/}} (v2.1,
\citealt{fm12}). We use 1000 walkers and obtain convergence\footnote{We checked 
for convergence of chains by examining the autocorrelation time of the chains
per dimension.} after a short burn-in phase of 100 steps per walker. The chains 
are then restarted around the best-fit value and evolved for another 4000 steps.
We find that the marginal posterior distributions of parameters are quite
Gaussian-like in shape, so we simply characterize their position using the
median, and their width using the central $68\%$ confidence limits (hereafter
CL; see Table~\ref{stream_parameters}).

@@

In principle, we could measure the proper motion of the Ophiuchus stream using
the proper motion of its confirmed members. However, since our sample of
confirmed members contains only 14 stars, there is a possibility that one or two
stars with incorrectly measured proper motions may bias the results. As an
example, stream member ``rgb4'' is clearly an outlier in proper motion as it has
$\mu_l\sim-30$ mas yr$^{-1}$, while the remaining members have $\mu_l\sim-6$ mas
yr$^{-1}$. A visual inspection of digitized photographic plates has revelead
that ``rgb4'' is blended with a neighbor of similar brightness, which affects
the measured position of the star and its proper motion.

@@

\begin{figure}
\plotone{orbit_XYZ.pdf}
\caption{
The orbit of the Ophiuchus stream in the past 370 Myr (about one orbital
period), shown in the galactocentric Cartesian coordinate system. Note the
pericenter passage at $t\approx-250$ Myr (solid square). Near this point in
time, the stream was also passing through the disk ($Z\sim0$ kpc, see the top
left panel) and was experiencing strong tidal forces due to disk shocking
(see Figure~\ref{tidal_force}).
\label{orbit_XYZ}}
\end{figure}

>>>>

\author{Branimir Sesar\altaffilmark{1,2}}
\author{Jo Bovy\altaffilmark{3,4}}
\author{Edouard J.~Bernard\altaffilmark{5}}
\author{Nelson Caldwell\altaffilmark{6}}
\author{Judith G.~Cohen\altaffilmark{7}}
\author{Morgan Fouesneau\altaffilmark{1}}
\author{Christian I.~Johnson\altaffilmark{6}}
\author{Melissa Ness\altaffilmark{1}}
%\author{Maria Bergemann\altaffilmark{1}}
\author{Annette M.~N.~Ferguson\altaffilmark{5}}
\author{Nicolas F.~Martin\altaffilmark{8,1}}
\author{Hans-Walter Rix\altaffilmark{1}}
\author{Edward F.~Schlafly\altaffilmark{1}}
%\author{builders}
\author{William S.~Burgett\altaffilmark{15}}
\author{Kenneth C.~Chambers\altaffilmark{9}} 
%\author{L.~Denneau\altaffilmark{9}} 
%\author{P.~Draper\altaffilmark{10}} 
\author{Heather Flewelling\altaffilmark{9}} 
%\author{T.~Grav\altaffilmark{11}} 
%\author{J.~N.~Heasley\altaffilmark{9}} 
\author{Klaus W.~Hodapp\altaffilmark{9}} 
%\author{M.~E.~Huber\altaffilmark{9}} 
%\author{R.~Jedicke\altaffilmark{9}} 
\author{Nick Kaiser\altaffilmark{9}} 
%\author{R.-P.~Kudritzki\altaffilmark{9}}
%\author{G.~A.~Luppino\altaffilmark{9}}
%\author{R.~H.~Lupton\altaffilmark{12}}
\author{Eugene A.~Magnier\altaffilmark{9}}
\author{Imants Platais\altaffilmark{11}}
%\author{N.~Metcalfe\altaffilmark{10}}
%\author{D.~G.~Monet\altaffilmark{13}}
%\author{J.~S.~Morgan\altaffilmark{9}}
%\author{P.~M.~Onaka\altaffilmark{9}}
%\author{P.~A.~Price\altaffilmark{12}}
%\author{C.~W.~Stubbs\altaffilmark{14}}
%\author{W.~Sweeney\altaffilmark{9}}
\author{John L.~Tonry\altaffilmark{9}}
%\author{R.~J.~Wainscoat\altaffilmark{9}}
\author{Christopher Waters\altaffilmark{9}}
\author{Rosemary~F.~G.~Wyse\altaffilmark{11}}

@@

We determined stellar parameters from the continuum-normalized, radial
velocity-corrected spectra using the SMH code of \citet{cas14}, which is built
on the MOOG code of \citet{sne73}. Kurucz $\alpha-$enhanced (0.4 dex) model
atmospheres \citep{ck04} and a line list compiled from \citet{fre10} by
\citet{cas14} were used (see Table~5 in the electronic version of the Journal).
First, effective temperatures were calculated from 2MASS photometry and
color-temperature calibrations of \citet{ghb09} and spectroscopic temperatures
were optimized around this value using the SMH code, by removing abundance
trends with line excitation potential. Parameters of $\log g$ and ${\rm [Fe/H]}$
were determined using 14 Fe I and 2-3 Fe II lines to achieve ionization balance,
and microturbulence was calculated by removing abundance trends with reduced
equivalent width of the lines. Estimates of the $\alpha-$enhancement were
obtained using only the few available clean Mg, Ca and Ti (I and II) lines,
which comprised a total of 6-8 lines per star. The abundances were measured from
equivalent widths and the lines we used are listed in Table~5. Because Mg lines
are strong and may be saturated, the values of ${\rm [Mg/Fe]}$ are significantly
different than values of ${\rm [Ca/Fe]}$ and ${\rm [Ti/Fe]}$.

@@

Having tied the positions for each object at one 2MASS and five USNO-B epochs to
the PS1 astrometric reference frame, we can now check for any additional
systematic uncertainties in the calibrated astrometry. We do so using the
leave-one-out cross-validation. One of the six calibrated positions is withheld,
and a straight line is fitted to the remaining five positions and the PS1
position. The straight line fit (i.e., essentially a proper motion fit,
neglecting the parallax) is then used to predict the position of an object at
the withheld epoch. The difference between the withheld position and the
predicted position is labeled as $\Delta RA$ or $\Delta Decl$.

@@

To efficiently explore the parameter space, we use the \citet{gw10} Affine
Invariant Markov chain Monte Carlo (MCMC) Ensemble sampler as implemented in the
{\em emcee} package\footnote{\url{http://dan.iel.fm/emcee/current/}} (v2.1,
\citealt{fm12}). We use 1000 walkers and obtain convergence\footnote{We checked 
for convergence of chains by examining the auto-correlation time of the chains
per dimension.} after a short burn-in phase of 100 steps per walker. The chains 
are then restarted around the best-fit value and evolved for another 4000 steps.
We find that the marginal posterior distributions of parameters are quite
Gaussian-like in shape, so we simply characterize their position using the
median, and their width using the central $68\%$ confidence limits (hereafter
CL; see Table~\ref{stream_parameters}).

@@

In principle, we could measure the proper motion of the Ophiuchus stream using
the proper motion of its confirmed members. However, since our sample of
confirmed members contains only 14 stars, there is a possibility that one or two
stars with incorrectly measured proper motions may bias the results. As an
example, stream member ``rgb4'' is clearly an outlier in proper motion as it has
$\mu_l\sim-30$ mas yr$^{-1}$, while the remaining members have $\mu_l\sim-6$ mas
yr$^{-1}$. A visual inspection of digitized photographic plates has revealed
that ``rgb4'' is blended with a neighbor of similar brightness, which affects
the measured position of the star and its proper motion.

@@

\begin{figure}
\plotone{orbit_XYZ.pdf}
\caption{
The orbit of the Ophiuchus stream in the past 370 Myr (about one orbital
period), shown in a right-handed galactocentric Cartesian coordinate system.
In this coordinate system, the Sun is at $(X,Y,Z)=(8,0,0)$ kpc and the y-axis is
positive towards galactic longitude $l=90\arcdeg$. Note the pericenter passage
at $t\approx-250$ Myr (solid square). Near this point in time, the stream was
also passing through the disk ($Z\sim0$ kpc, see the top left panel) and was 
experiencing strong tidal forces due to disk shocking (see
Figure~\ref{tidal_force}).
\label{orbit_XYZ}}
\end{figure}
